---
title: "Data Wrangling L1"
output: github_document
---

Let's import the FAS_litters.csv using a relative path.

```{r setup}
library(tidyverse)
```


```{r}
liters_df=
  read_csv("data/FAS_litters.csv")

liters_df = 
  janitor::clean_names(liters_df)
```

Let's import the FAS_litters.csv using an absolute path. Having issues with this, maybe bec of spaces in file names.

```{r, eval=FALSE}
liters_df_abs=
  read_csv("~Desktop/Fall2023/DataScience/Lectures/DataWrangling/Data/FAS_litters.csv")

liters_df_abs = 
  janitor::clean_names(liters_df_abs)

head(liters_df)

tail(liters_df)
```

```{r, eval=FALSE}
View(liters_df)
```

Look at data summary

```{r}
str(liters_df)

skimr::skim(liters_df)

```

Import pups dataset
```{r}
pups_df=
  read_csv("data/FAS_pups.csv")

pups_df = 
  janitor::clean_names(pups_df)
```

#Options in 'read_*'

Col_names - assignes first row as names. And Skip skips rows.
```{r}
liters_df
  read_csv("data/FAS_litters.csv",
  skip = 10, col_names = FALSE)
```

Look at NA values.
Here we say treat NA and 19 as missing.
```{r}
liters_df
  read_csv("data/FAS_litters.csv",
  na= c("NA", 19 ))
```

Column types
Change variables from numeric to character to factor etc.

```{r}
liters_df
  read_csv("data/FAS_litters.csv",
  col_types = 
    cols(
      Group = col_factor()
    ))
```

#Other file types

Import an xlx file first. 
First load in readdxl library

```{r}
  library(readxl)
```

```{r}
  mlb_df=
    read_excel("data/mlb11.xlsx")
```

Import a sas file.

```{r}
  library(haven)
```

```{r}
  pulse_df=
    read_sas("data/public_pulse_data.sas7bdat")
```

```{r, eval=FALSE}
View(pulse_df)
```


#Base R

```{r, eval= FALSE}
liters_df = 
  read.csv("data/FAS_litters.csv")

liters_df$Gr
```

Dont use $ because it calls in all variables with the same beginning if you dont type it out full name.

#Export Data

We want to export data.
```{r}
liters_df_cleaned=
  read_csv("data/FAS_litters.csv")

liters_df_cleaned = 
  janitor:: clean_names(liters_df_cleaned)

write_csv(liters_df_cleaned, "data/liters_cleaned.csv")
```