---
title: "Data manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

Import FAS litters and pups
```{r import}

litters_df = 
  read.csv("data/FAS_litters.csv")
litters_df= 
  janitor:: clean_names(litters_df)

pups_df = 
  read.csv("data/FAS_pups.csv")
pups_df= 
  janitor:: clean_names(pups_df)
```

#select

Select is used to select columns. Can change order with this. The ":" means everything between variables. 

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, gd0_weight, group)

select(litters_df, group, gd0_weight:gd_of_birth)
```

Starts with includes all variables that start with certain letters, can also do ends with and contains. 

```{r}
select(litters_df, gd0_weight, starts_with("pups"))
```

can also tell it what to remove (-).
```{r}
select(litters_df, - gd0_weight, -starts_with("pups"))
```


Can also rename during select. But can rename without select and keep everything.
```{r}
select(litters_df, group, litter_id = litter_number)

rename(litters_df, litter_id = litter_number)
```

Can also keep everything and reorder.

```{r}
select(litters_df, litter_number, everything())
```

Can also keep reorder.

```{r}
relocate(litters_df, litter_number)
```

select vs pull. Select keeps the dataframe and pull creates a vector (gets something out of data frame. 

```{r}
select(litters_df, group)

pull(litters_df, group)
```

#filter

Get rid of rows using filter. =1 is not equal to.

```{r}
filter(litters_df, group=="Con7")

filter(litters_df, gd0_weight>20)
```

keep this or that
```{r}
filter(litters_df, group %in% c("Con7","Con8"))
```

Dont keep this (keep everything else)
```{r}
filter(litters_df, !(group=="Con7"))
```

Select multiple things at once.
```{r}
filter(litters_df, group %in% c("Con7","Con8"), gd0_weight>20)
```

what about missing rows
```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

#mutate

This is used to add or change variables.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```

##arange 
Puts in order.

```{r}
arrange(litters_df, gd0_weight)
```

#pipe

```{r}
litters_df
  read_csv("data/FAS_litters.csv") |> 
    janitor:: clean_names() |>
    select(-starts_with("pups")) |>
  mutate(
    group= str_to_lower(group),
    weight_gain = gd18_weight - gd0_weight
  ) |>
    drop_na(weight_gain) |>
    arrange(group, weight_gain)
```

